IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-10-17 paper/retrofit.tex
improve figures for one-column layout

\paragraph{Exceptions versus message passing}
Another language feature that has caused difficulty for
many information flow control systems is exception handling
when built on top of this stack discipline.
%
The key difficulty appears when considering the propagation of exceptions
across stack boundaries where the current label is restored.
%
If the exception crosses this boundary, it can leak information about the
inner secret computation into a public computation; hence, one must
explicitly ensure that |toLabeled| does not propagate exceptions.

When stack frames are treated as isolated tasks, the correct semantics
are achieved automatically.
%
Exceptions simply stop when the hit the top-level of a task:
the only way for the frame to communicate information is by
initiating inter-task message passing.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.

@@

\begin{figure}
\begin{code}
tv   ::= \tx.te | true | false | ta
te   ::= tv | tx | te te | if te then te else te
       | ref te | !te | te := te | fix te
tE   ::= tHole | tE te | tv tE | if tE then te else te
       | ref tE | !tE | tE := te | tv := tE | fix tE

te1 ; te2            @= (\tx.te2) te1  where  tx notmember FV(te2)
let tx = te1 in te2  @= (\tx.te2) te1
\end{code}

\begin{mathpar}

@@

\begin{figure}
\begin{code}
lop  ::=  canFlowTo | lub | glb
iv   ::=  id | il | true | false | unit
ie   ::=  iv | ix | ie lop ie | getLabel | setLabel ie | getTaskId
       |  sandbox ie | send ie ie ie | recvf ix ix ie ie
iE   ::=  iHole | iE lop ie | iv lop iE | setLabel iE
       |  send iE ie ie | send iv iE ie | send iv iv iE
io   ::=  (il, id ie)
Q    ::=  nil | io , Q
iS   ::=  emptyset | bracket iS (mapsto id Q)
\end{code}

\begin{mathpar}

@@

\inferrule[I-labelOp]
{ |denot (il1 lop il2) = iv|}
{ |niconf iS id il (il1 lop il2) -> niconf iS id il iv| }
\end{mathpar}
\caption{IFC language with all single-task operations.}
\label{fig:ifc}
\end{figure}

@@

\begin{figure}
%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
%format .-/>  = "\not\overset{\alpha}{\hookrightarrow}"
\begin{tabular}{ll}
\begin{minipage}{.22\textwidth}
\begin{code}
iv  ::= cdots | IT tv
ie  ::= cdots | IT te
iE  ::= cdots | IT tE
\end{code}
\end{minipage} &
\begin{minipage}{.22\textwidth}
\begin{code}
tv  ::= cdots | TI iv
te  ::= cdots | TI ie
tE  ::= cdots | TI iE
\end{code}
\end{minipage}
\end{tabular}

\begin{code}
conf tS te @= iconf iS (fullconf id il tS (iniEt te), ldots)
niconf iS id il ie @= iconf iS (fullconf id il tS (iniEi ie), ldots)
CE[e] -> iconf iS (it, ldots) @= CE[e] .-> iconf iS (sched step (it, ldots))
\end{code}

\begin{mathpar}
\inferrule[I-sandbox]
{ 
|iS' = bracket iS (mapsto id' nil)|\\
|tS' = klone(tS)|\\
|it1 = fullconf id il tS (iniE id')|\\
|itnew = fullconf id' il tS' ie|\\
|fresh id'|
}
{|
iconf iS (fullconf id il tS (iniEi (sandbox ie)), ldots)
.->
iS'; sched box (it1, ldots, itnew)
|}

@@

\begin{figure}
  Concurrent, round robin:
\begin{code}
  roundrobin step (it1, it2, ldots)    = it2, ldots, it1
  roundrobin done (it1, it2, ldots)    = it2, ldots
  roundrobin nostep (it1, it2, ldots)  = it2, ldots
  roundrobin box (it1, it2, ldots)     = it2, ldots, it1
\end{code}
  Sequential:
\begin{code}
  seq step (it1, it2, ldots)      = it1, it2, ldots
  seq nostep (it1, it2, ldots)    = it1, it2, ldots
  seq done (it)                   = it
  seq done (it1, it2, ldots)      = it2, ldots
  seq box (it1, it2, ldots, itn)  = itn, it1, it2, ldots
\end{code}
\caption{Scheduling policies}
\label{fig:scheduler}
\end{figure}

>>>>

\paragraph{Exceptions versus message passing}
Another language feature that has caused difficulty for
many information flow control systems is exception handling
when built on top of this stack discipline.
%
The key difficulty appears when considering the propagation of exceptions
across stack boundaries where the current label is restored.
%
If the exception crosses this boundary, it can leak information about the
inner secret computation into a public computation; hence, one must
explicitly ensure that |toLabeled| does not propagate exceptions.
%
When stack frames are treated as isolated tasks, the correct semantics
are achieved automatically.
%
Exceptions simply stop when the hit the top-level of a task:
the only way for the frame to communicate information is by
initiating inter-task message passing.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.

@@

\begin{figure}
\begin{code}
tv   ::= \tx.te | true | false | ta
te   ::= tv | tx | te te | if te then te else te | ref te | !te | te := te | fix te
tE   ::= tHole | tE te | tv tE | if tE then te else te | ref tE | !tE | tE := te | tv := tE | fix tE
te1 ; te2            @= (\tx.te2) te1  where  tx notmember FV(te2)
let tx = te1 in te2  @= (\tx.te2) te1
\end{code}
\begin{mathpar}

@@

\begin{figure}
\begin{code}
iv   ::=  id | il | true | false | unit qquad qquad lop  ::=  canFlowTo | lub | glb
ie   ::=  iv | ix | ie lop ie | getLabel | setLabel ie | getTaskId |  sandbox ie
       | send ie ie ie | recvf ix ix ie ie
iE   ::=  iHole | iE lop ie | iv lop iE | setLabel iE |  send iE ie ie | send iv iE ie | send iv iv iE
io   ::=  (il, id ie)  qquad qquad   Q    ::=  nil | io , Q   qquad   iS   ::=  emptyset | bracket iS (mapsto id Q)
\end{code}
\begin{mathpar}

@@

\inferrule[I-labelOp]
{ |denot (il1 lop il2) = iv|}
{ |niconf iS id il (il1 lop il2) -> niconf iS id il iv| }

\end{mathpar}
\caption{IFC language with all single-task operations.}
\label{fig:ifc}
\end{figure}

@@

\begin{figure}
%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
%format .-/>  = "\not\overset{\alpha}{\hookrightarrow}"
\begin{tabular}{lll}
\begin{minipage}{.22\textwidth}
\begin{code}
iv  ::= cdots | IT tv
ie  ::= cdots | IT te
iE  ::= cdots | IT tE
\end{code}
\end{minipage} &
\begin{minipage}{.22\textwidth}
\begin{code}
tv  ::= cdots | TI iv
te  ::= cdots | TI ie
tE  ::= cdots | TI iE
\end{code}
\end{minipage} &
\begin{minipage}{.42\textwidth}
\begin{code}
conf tS te @= iconf iS (fullconf id il tS (iniEt te), ldots)
niconf iS id il ie @= iconf iS (fullconf id il tS (iniEi ie), ldots)
CE[e] -> iconf iS (it, ldots) @= CE[e] .-> iconf iS (sched step (it, ldots))
\end{code}
\end{minipage}
\end{tabular}
\begin{mathpar}
\inferrule[I-sandbox]
{ 
|iS' = bracket iS (mapsto id' nil)|\\
|tS' = klone(tS)|\\
|it1 = fullconf id il tS (iniE id')|\\
|itnew = fullconf id' il tS' ie|\\
|fresh id'|
}
{|
iconf iS (fullconf id il tS (iniEi (sandbox ie)), ldots)
.->
iS'; sched box (it1, ldots, itnew)
|}

@@

\begin{figure}
\begin{tabular}{ll}
\begin{minipage}{.45\textwidth}
\begin{code}
  roundrobin step (it1, it2, ldots)    = it2, ldots, it1
  roundrobin done (it1, it2, ldots)    = it2, ldots
  roundrobin nostep (it1, it2, ldots)  = it2, ldots
  roundrobin box (it1, it2, ldots)     = it2, ldots, it1
\end{code}
\end{minipage} &
\begin{minipage}{.45\textwidth}
\begin{code}
  seq step (it1, it2, ldots)      = it1, it2, ldots
  seq nostep (it1, it2, ldots)    = it1, it2, ldots
  seq done (it)                   = it
  seq done (it1, it2, ldots)      = it2, ldots
  seq box (it1, it2, ldots, itn)  = itn, it1, it2, ldots
\end{code}
\end{minipage}
\end{tabular}
\caption{Scheduling policies (concurrent round robin on the left, sequential on the right).}
\label{fig:scheduler}
\end{figure}
